---
title: "STAT 424 HW1"
subtitle: Fall 2020
output:
  pdf_document: default
  word_document: default
---

## Instructions

1. Submit your solution on Canvas as a pdf file before *September 11 at 11:59pm*.
2. Use R for all computation, and include your code. 
3. We recommend you start with the template `hw1.Rmd` file on Canvas.

## Rubric

* [30 points] One problem below will be graded for,
  - [15 points] Correctness. Are all parts answered correctly?
  - [15 points] Clarity. Are all answers justified? For questions with a coding
  component, is code provided, and is it readable?
* [20 = 4 * 5 points] Four problems below will be graded on completeness.
* Total: 50 points

```{r setup, include=FALSE}
library("tibble")
library(ggplot2)
library("broom")
library("readr")
library("reshape2")
knitr::opts_chunk$set(echo = TRUE)
```

## Problems

**1. [DAE 1.3]** Suppose that you want to compare the growth of garden flowers
with different conditions of sunlight, water, fertilizer, and soil conditions.
Complete steps 1 - 3 of the guidelines for designing experiments in Section 1.4.
This is an open-ended question -- there is no one right answer -- just explain
your thought process.

1. Recognition of and statement of the problem: The problem is the potential difference in growth of flowers with different conditions.

2. Selection of the response variable: The response variable in this problem should be the growth height.

3. Choice of factors, levels, and range: Controllable Factors -- plant type, fertilizer type, pot/location of planting. Uncontrollable factors -- sunshine, rain levels, air conditions. Blocking factors -- I guess you could split each existing plant in half since they would likely react the same to the same conditions.

**2. [DAE 1.8 / MED 1.5]**

  a. What is replication? Why do we need replication in an
  experiment? 
  
  **Replication is when an experiment with the same conditions is repeated in order to see variability and make sure our conclusions are reproducable.**
  
  b. Present an example that illustrates the difference between
  replication and repeated measurements. 
  
  **Replication is where you recreate an experiment several times, with different plants. Repeated measurements would be measuring the same plants several times.**
  
  c. State whether or not replication in your example would have any
  nonstatistical disadvantages, such as a considerable increase in the cost of running the experiment.
  
  **There would be several disadvantages such as the cost of buying a lot of plants, the cost of time, space, and other resources the plants take up. Personally, my family gets annoyed when I bring home more plants since I already have ~too many~.**

**3. [DAE 1.9, modified]**

  a. Why is randomization important in an experiment?
  
  **It is important in order to limit certain biases that could arise from testing things in a non random order.**
  
  b. Give one example where complete randomization would be difficult or unethical^[The field of observational causal inference exists for drawing conclusions in this case, but it is out of scope for our class.].
  
  **Testing negative side effects of bad behaviors or medicines would be unethical to randomize since you would intentionally subject participants to harm.**

**4. [MED, 1.4, modified]** An experiment with a single factor and two levels, 1 and 2, was used, and the results were as follows:

```{r}
df <- tibble(
  "sample_id" = as.factor(seq(10)),
  "1" = c(6.1, 8.2, 7.3, 8.4, 8.0, 7.6, 8.7, 9.3, 6.8, 7.5),
  "2" = c(6.3, 8.0, 7.7, 8.1, 8.6, 7.2, 8.4, 9.7, 6.8, 7.2)
)
```

  a. Make two histograms, one for when factor 1 was used, another for when factor
  2 was used.
  
```{r}
hist(df$`1`)
```
```{r}
hist(df$`2`)
```

  
  b. Plot the data with the sample ID as the $x$-axis, the measured value as the
  $y$-axis, and the factor level as the color of the point.
  

```{r}
new.df <- melt(df, measure.vars = c("1", "2"), variable.name = "Factors", value.name = "strength")

g <- ggplot(new.df, aes(x=sample_id, y=strength))
g + geom_point(data = new.df[new.df$Factors == 1, ], colour = "red") + geom_point(data = new.df[new.df$Factors == 2, ], colour = "blue")
```

  
  c. What does part (b) tell you about the need for blocking? Is it needed -- why or why not?
  
**This would be a good experiment for blocking because the intervals between factor 1 and factor 2 for each sample ID are similar, but it varies as to which factor is on top. This either means that there is not much of a difference between factor 1 and factor 2, or there is need for more measures to block out any unaccounted for environmental effects.**
    
**5. [MED 1.15]** Assume that an experiment with four levels of a single factor was run, and the value of the response variable was almost strictly increasing during the runs of the experiment. What would you suspect and what would be your recommendation?

**I would think there are environmental factors at play that were not controlled for between runs. I would recommend adding more control factors until you can find out what the confounding variable is.**

## Feedback

a. How much time did you spend on this homework?

**about an hour and a half**

b. Which problem did you find most valuable?
**I thought the graphing one was nice since I haven't had to visualize data in a while and I need some refreshers on the code. I also thought it was a confusing problem since it is not clear if the experiment only has 10 subjects that each undergo treatment twice of if there are 20 total subjects.**


## References

[DAE] Montgomery, Douglas C. Design and analysis of experiments. John Wiley &
sons, 2017.

[MED] Ryan, Thomas P. Modern experimental design. Hoboken, N.J:
Wiley-Interscience, 2007. Print. Accessed for free at
https://search.library.wisc.edu/catalog/9912216524202121.
